A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2019-01-27 article/05_Conclusion.Rmd
more work on the conclusion

The major questions posed in the introduction of this paper (section \ref{introduction}) have been answered (section \ref{discussion}): *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*. Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information. Although only `r r("gpr size", sf_desc)` graves of a six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research. The time series can be employed for cultural distance calculation and they yield the surprising observation, that cultural distance and spatial distance do not significantly correlate. Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.

@@

1. For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour reconstruction in archaeology. The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives. Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances archaeological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].

2. test

>>>>

The major questions posed in the introduction of this paper have been answered : *Bulk radiocarbon data is a powerful source to trace general trends in prehistoric Europe* and *allows long-term observations concerning cultural distance*. Radiocarbon dates can be transformed to meaningful regional time series, if they can be linked to the relevant meta information. Although only `r r("gpr size", sf_desc)` graves of a maybe six-figure statistical population that can be estimated for all burials in the European Bronze Age [@nikulkaArchaologischeDemographieMethoden2016] have been available for this analysis, the resulting reconstruction of general burial tradition trends in wast areas of Central, Northern and North-western Europe comes close to the one achieved by literature research. The time series can be employed for cultural distance calculation and they yield the surprising observation that cultural distance and spatial distance do not significantly correlate. Although this interpretation must be treated with caution, it seems as if the spread of burial traditions are not simple *Diffusion of Innovation* events as introduced by @rogersDiffusionInnovations1962, but the result of more complex interactions between competing ideas.

Future research is required to de-tangle these interactions -- of course on the basis of more precise data. This endeavour will not end with the simple description of social learning cascades, but with the analysis of the initial reasons for diversification. From the point of view of Cultural Complexity -- the question, which factors accelerate or decelerate the variability of human behaviour [@henrichDemographyCulturalEvolution2004; @mesoudiWhatCumulativeCultural2018; @vaesenPopulationSizeDoes2016] -- the long term fluctuations of dominance of one burial rite over the other are remarkable. Its standard explanation models (among others population density and environmental risk) can not be applied to stylistic variables.

@@

1. For this paper burial traditions were the main object of research, but the same methodological approach could be applied for other aspects of cultural behaviour reconstruction in archaeology. The idea to systematically collect data points that combine spatial and temporal location with simplified meta information about cultural phenomena might be a way to achieve better reproducibility of overarching narratives. Fast and easy *reproducibility* is of crucial importance to achieve scientific progress and it includes the notion of rather technical *preproducibility* [@starkReproducibilityMustCome2018]: Data (and code) must be shared to fuel advances in anthropological theory building [@kitzesPracticeReproducibleResearch2017; @marwickComputationalReproducibilityArchaeological2017].

2. This paper is built upon Cultural Evolution theory (see section \ref{introduction}). It not just relies on terminology and methodology developed in this field, the process of abstraction necessary to go from radiocarbon dates to cultural distance was also heavily influenced by Evolutionary archaeology. This theoretical framework might also be a source of inspiration and methods for other projects. The adoption and adaption of statistical tools and models from population genetics into anthropology was and is a feature of Cultural Evolution theory [@creanzaCulturalEvolutionaryTheory2017].
